Two Mode Networks and Relational Similarity

### 2023-09-08 main.tex

Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities. This applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.

I also show that CA can be interpreted as providing \textit{positional} information on people and groups. First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar patterns of connectivity to other similar entities in the other mode (e.g., groups). Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode. Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.

\subsection{Organization of the Paper}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via singular value decomposition (SVD) of the (inverse degree-weighted) one-mode projection of the affiliation matrix. This approach is computationally and mathematically related but \textit{not} equivalent to the CA of the affiliation matrix, which results in the usual Bonacich scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Reflective Centralities in Two-Mode Networks}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as ``economic complexity.'' Subsequent shows that there is no logical connection between the formal method and this particular application. As such, I introduce it here using the more intuitive---and classical from a social science perspective---case of the duality of persons and groups \citep{breiger1974duality}.

@@

That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups. As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.

@@

Equation \ref{eq:R1_p} says ``people are more central when the average sum of the memberships of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation \ref{eq:R2_g} says ``groups are more central when the average sum of the number of memberships of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number). Of course, we can keep on going and define third-order reflections.

@@

Equation \ref{eq:R3_p} says something like ``people are more central when the average of the average number of memberships of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation \ref{eq:R3_g} says, ''groups are more central when the average number of members of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).

Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, their degree centrality). On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to. In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their degree centrality). On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{groups their members belong to}. While these are distinct metrics, in practice, the order of nodes in each mode ends up being identical when the ranks ``freeze'' after some number of iterations (proportional to the network size).

@@

Equation~\ref{eq:Rq_p} says that the reflective centrality of person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$. Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.

@@

\subsection{Empirical Example}
Figure~\ref{fig:pg-reflections} shows the trajectory of the Hidalgo-Hausmann reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}. The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

@@

Where $A$ is the original affiliation matrix. The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors of the affiliation matrix corresponding to the largest eigenvalue $\lambda$, as given by the usual SVD of the affiliation matrix. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

@@

These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members. These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.

As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue problem. To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}. Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:

@@

Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell. In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell. It can be shown \citep{van2021correspondence}, that in the limit, the iterative Hidalgo-Hausmann \citeyearpar{hidalgo2009building} reflective centralities can be obtained as the solution of the eigensystem:

@@

Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups. The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode, essentially weighting the projection by node degree before performing the SVD.

@@

In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share a membership. Summed across groups $g$, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}. As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the larger eigenvalue of this matrix of shared memberships (for people) and shared people (for groups). The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.

@@

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical. In the Bonacich case, SVD of the one-mode projection of the two-mode network; in the reflective case, SVD of the degree-weighted one-mode projection of the same network. Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to. The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-sized adjusted} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.

@@

Figure~\ref{fig:ca-v-bon} illustrates the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}, which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis. If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position. Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach. This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.

@@

To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships). The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups). As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$. This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition. These are shown as inset plots in Figure~\ref{fig:eigvec} for both people and groups. Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.

@@

To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups). In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity. Also, the rows of each matrix are sorted according to $C^R$ for both people and groups. As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.

The bottom panel of Figure~\ref{fig:ca-sim} shows the correspondence plot of the first two CA dimensions. People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices. The plot reveals that the usual correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---is best thought of as a low-dimensional representation of the (other mode's) degree-weighted similarity network across people and groups. People with similar values in $S_p$ appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm), as do events with similar values in $S_g$.

Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to imply (for instance) that two people attend the same set of events, or that two groups share the same set of members. What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar after taking people's activity levels and the sizes of the groups they belong to into account. Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.

@@

This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$. But as we have seen, this is \textit{not} what CA is designed to do. Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes. Oddly, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two dimensions of the SVD of this matrix.

>>>>

Particularly, I show that CA can be thought of as computing a kind of \textit{dual centrality} score on both modes---one related to but not mathematically equivalent to the Bonacich \citeyearpar{bonacich1991simultaneous} two-mode centrality----an approach that, as noted by \citet{van2021correspondence}, has been recently re-invented in some corners of network science under the guise of the ``economic complexity index'' \citep{hidalgo2009building, mealy2019interpreting}. With \citet{van2021correspondence}, I argue that this is a rather restricted interpretation of a more general ``dual centrality'' score that connects to the use of CA in some disciplines (e.g., ecology) as a method of \textit{ordination}---e.g., the discovery of latent one-dimensional orderings among a set of entities. In social network analysis, this applies to conceptualizing the venerable connection between people and the groups they join \citep{breiger1974duality}.

Additionally, I argue that CA can also be used to extract latent \textit{positional} information on people and groups. First, I show that there is a suggestive similarity between the scores obtained from the first dimension of the CA of the two-mode affiliation matrix (the aforementioned dual centrality score) and recent work on ``generalized relational similarity'' in two-mode networks \citep{kovacs2010generalized, lizardo2024two}. As such, CA recovers clusters of entities (e.g., people) linked by their similar connectivity patterns to similar entities in the other mode (e.g., groups). Second, drawing on recent work by \citet{van2021correspondence}, I show that the first \textit{two} CA dimensions provide information on the degree-weighted similarity of nodes in the same mode concerning their connections to nodes in the other mode. Thus, CA can be used as an approach to indirect ``community'' identification in two-mode networks that connects to the use of CA in some disciplines (e.g., computer and information science) as a \textit{clustering} method \citep{zha2001bipartite}---e.g., the discovery of sociometrically similar subsets of entities in affiliation networks.

\subsection{Organization of the Paper}
The rest of the paper is organized as follows. In the next section, I motivate the use of CA as producing a type of Bonacich-style dual centrality via weighted iteration across the two modes of the affiliation matrix. This leads naturally to an abbreviated way of computing the same scores via eigenvector decomposition of the (inverse degree-weighted) one-mode projection of the affiliation matrix for each set of nodes. This approach is computationally and mathematically related but \textit{not} equivalent to the matrix factorization of the one-mode projection matrices, which results in the usual Bonacich dual centrality scores. I then draw on recent work by \citet{van2021correspondence} to show the links between the ordering of nodes along the first CA dimension and Kovacs's generalized relational similarity, and the distance of nodes in the space formed by the first two dimensions and the connectivity-similarity between nodes across modes. I contrast this ordering to that provided by the Bonacich dual centrality scores, which in contrast to the CA ordering---which reveals a dual community partition--- recovers instead the core-periphery structure of the two-mode network. In closing, I provide some pointers on how CA can be used as a method for structural (e.g., centrality and positional) analysis, independently of the usual emphasis on joint graphical displays and visual exploration.

\section{Reflective Centralities in Two-Mode Networks}
In a highly cited piece, \citet{hidalgo2009building} motivated what they saw at the time as a novel way of ranking nodes in a two-mode network---effectively computing a version of two-mode centrality---using what they called at the time a ``reflective'' approach. Hidalgo and Hausmann's original empirical application was to the two-mode country-by-product networks, hence the original baptizing of their approach as one geared to extracting the ``economic complexity'' of countries in the world system (and dually the complexity of given products). Subsequent shows that there is no logical connection between the formal method and this particular application. As such, I introduce it here using the more intuitive---and classical from a social network analysis perspective---case of the duality of persons and groups \citep{breiger1974duality}.

@@

That is, the first-order centrality of people is the row sum of the entries in the affiliation matrix $A$, and the column sums of the same matrix give the first-order centrality of groups. In what follows, I will generally refer to $C^R_p(1)$ as (individual) ``activity'' and $C^R_p(1)$ as (group) ``size.''

As noted by \citet{hidalgo2009building}, the key to the reflective approach is the observation that, once we have these first-order quantities, it is possible to compute ``second-order centralities'' $C^R(2)$ for both people and groups using the (averaged) first-order centralities of the entities in the other mode they are connected to.

@@

Equation \ref{eq:R2_p} says ``people are more central when the average sum of the size of the groups they belong to is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_g(1)$ is a big number). Equation \ref{eq:R2_g} says ``groups are more central when the average activity of their members is large'' (e.g., whenever $a_{pg} = 1$ and $C^R_p(1)$ is a big number). Of course, we can keep on going and define third-order reflections.

@@

Equation \ref{eq:R3_p} says something like ``people are more central when the average sum of the average activity of the members of the groups they belong to is large'' (e.g., $a_{pg} = 1$ and $C^R_g(2)$ is a big number). Equation \ref{eq:R3_g} says, ''groups are more central when the average sum of the average size of the groups their members belong to is large.'' (e.g., $a_{pg} = 1$ and $C^R_p(2)$ is a big number).

Note that for the people, the even-numbered reflection $C^R_p(2)$ assigns centrality based on a formal feature of the \textit{groups} they belong to (in this case, the group sizes). On the other hand, the odd-numbered reflection $C^R_p(3)$ assigns centrality based on a formal feature of the \textit{members of the groups} they belong to (in this case, the average size of the groups they belong to). In the same way, for the groups, the even-numbered reflection $C^R_g(2)$ assigns centrality based on a formal feature of the \textit{people} who belong to them (in this case, their activity). On the other hand, the odd-numbered reflection $C^R_g(3)$ assigns centrality based on a formal feature of the \textit{other groups their members belong to} (in this case, their average group size). While these are distinct metrics, in practice, the ordering of the nodes in each mode ends up being identical across even and odd centralities after the ranks ``freeze'' past a given number of iterations (proportional to the network size).

@@

Equation~\ref{eq:Rq_p} says that the reflective centrality of a person \textit{p} at iteration \textit{q} is the sum of the $q-1$ centralities of the groups they belong to ($C^R_g(q-1)$) divided by their number of memberships $C^{R}_p(1)$. Equation~\ref{eq:Rq_g} says that the $q^{th}$ group reflective centrality is the sum of $q-1$ centralities $C^R_p(q-1)$ of their members, divided by the number of group members $C^R_g(1)$.

@@

\subsection{Empirical Example}
Figure~\ref{fig:pg-reflections} shows the trajectory of the HH reflective centralities for persons (top panel) and groups (bottom panel) in the Southern Women data \citep{davis1941}, shown in Table~\ref{tab:southern}. The rank order of people and groups in the $q^{th}$ centrality is plotted in the y-axis, and the centrality iteration $q$ is plotted on the x-axis.

@@

Where $A$ is the original affiliation matrix. The unknown $C^B_g$ and $C^B_p$ scores can be obtained from the row and column eigenvectors corresponding to the largest eigenvalue $\lambda$, as given by the singular value decomposition (SVD) of the rectangular affiliation matrix. Importantly, as Faust \citeyearpar[170]{faust1997centrality} notes, it is possible to express Equations~\ref{eq:bon1} and~\ref{eq:bon2} in terms of person and group-specific centralities.

@@

Where $\lambda$ is the (first) eigenvalue corresponding to the eigenvector containing the $C^B_p$ scores. These equations make clear that in the Bonacich two-mode centrality of people is proportional to centralities of the groups they join, and the centrality of groups is proportional to the centralities of their members. These capture the duality property since the centralities of nodes in each mode are given by aggregating their connections to nodes in the other mode. Note the formal similarity between equations~\ref{eq:faust1} and~\ref{eq:faust2} and equations~\ref{eq:Rq_p} and~\ref{eq:Rq_g}. The main difference is that in the reflective equations, a person's centrality is proportional to the \textit{activity-weighted} centralities of the groups they join, and a group's centrality is proportional to the \textit{group size weighted} centrality of the people who are members.

As noted by \citet{mealy2019interpreting} and \citet{van2021correspondence}, Hidalgo and Hausmann's \citeyearpar{hidalgo2009building} reflective centralities can also be obtained as the solution to an eigenvalue (matrix factorization) problem. To see this, recall that, as Bonacich \citeyearpar[157]{bonacich1991simultaneous} notes, we can solve for $C^B_g$ in~\ref{eq:bon2} and $C^B_p$ in~\ref{eq:bon1} and substitute the respective solutions into \ref{eq:bon1} and~\ref{eq:bon2}. Matrix-algebraic reduction of the resulting equations would show that the Bonacich dual centralities can also be obtained as solutions to the eigensystem:

@@

This shows that the dual centrality Bonacich scores for persons and groups are equivalent to the eigenvectors of the respective one-mode projection matrices corresponding to the first (largest) eigenvalue.

Now, consider the $|P| \times |P|$ matrix $Dp$, which contains the ``first order'' reflective centralities of each person $C^R(1)_p$ (activity) along the diagonals and zeroes in every other cell. In the same way, consider the $|G| \times |G|$ matrix $Dg$, which contains the ``first order'' (degree) centralities of each group $C^R(1)_g$ (size) along the diagonals and zeroes in every other cell. It can be shown \citep{van2021correspondence}, that in the limit ($q \rightarrow \infty$), the iterative HH reflective centralities can be obtained as the solution of the eigensystem:

@@

Note the formal similarity (and key differences) between these equations and the Bonacich two-mode centralities in equations~\ref{eq:bon3} and~\ref{eq:bon4}. Both extract individual and group centralities as eigenvectors of the one-mode projection of the original affiliation matrix: $AA^T$ for people and $A^TA$ for groups. The difference is that the reflective centralities pre-multiply the affiliation matrix and its transpose by the inverse of the first-order centralities of the nodes in each mode before factorizing, essentially normalizing the one-mode projection by the degrees of both sets of nodes.

@@

In equation~\ref{eq:mealy1}, the numerator is equal to one when person $p$ and person $p'$ share membership in a group $g$. Summed across groups, this gives the number of memberships that $p$ and $p'$ have in common \citep{breiger1974duality}. As noted, the Bonacich dual centralities are obtained from the eigenvector corresponding to the first eigenvalue of this matrix of shared memberships (for people) and shared people (for groups). The reflective centralities, on the other hand, are given by the eigenvectors of a weighted version of the same matrix, where the weights are the sizes of each of the groups $p$ shares with each other person summed across groups and divided by the total number of $p$'s memberships.

@@

Accordingly, it would make no sense to call the Bonacich scores $C^B_p$ and $C^B_g$ from equations~\ref{eq:bon3} and~\ref{eq:bon4} ``centrality scores'' but fail to hold that designation from $C^R_p$ and $C^R_g$, given that the mathematics are not only ``similar'' \citep[162]{bonacich1991simultaneous} but formally identical. In the Bonacich case, a factorization of the one-mode projection of the two-mode network; in the reflective case, factorization of the degree-weighted one-mode projection of the same network. Accordingly, the CA of a two-mode network will return---along the first dimension---a rank-ordered score for nodes in each mode that meets all the conditions for qualifying as a centrality measure for two-mode networks. Particularly, the CA centrality retains the duality property \citep[128]{faust2005using}, as the (average) centrality of people is a function of the (average) centrality of the groups they belong to. The (average) centrality of groups, in turn, is a function of their members' (average) centrality. The person and group score ranks given by the first CA dimension are thus better thought of as \textit{activity and group-size normalized} versions of the familiar Bonacich \citeyearpar{bonacich1991simultaneous} dual centralities for two-mode networks.

@@

\section{Illustrative Analysis of the Southern Women Data}
\subsection{Row and Column Re-ordered Affiliation Matrices}
Figures~\ref{fig:ca-v-bon}(a) and~\ref{fig:ca-v-bon}(b) illustrate the key differences between the two versions of dual centrality $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension in (a) and the \citet{bonacich1991simultaneous} centrality in (b). In the plot, a cell entry is colored red if it has a one in the original affiliation matrix and is white when the corresponding entry is zero. As we can see, the two row-column-reshuffled matrices have appreciably distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a dual community partition between groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward community partitioning and $C^B$ more focused on finding a core of well-connected people and groups.

\subsection{Eigenvector Plot}
We can confirm that the first CA axis points toward the underlying community partitioning of the two-mode network by looking at Figure~\ref{fig:eigvec}(a) and~\ref{fig:eigvec}(c), which shows the $C^R$ score of persons and groups on the x-axis against the score's rank order on the y-axis. If a two-mode network has no discernible community structure, the first dimension CA scores would be distributed as a continuous logistic curve; when community structure is present, we should observe discernible breaks in this distribution \citep{van2021correspondence}. The Southern Women data clearly belong to the second category. In the people mode, we have $\left[Frances, Laura, Brenda, Charlotte, Evenly, Theresa, Elanor, Ruth, Pearl\right]$ on one side, $\left[Dorothy, Nora, Katherine, Sylvia, Flora, Olivia, Myra, Helen, Verne\right]$ on the other. Among events, we have $\left[E1, E2, E3, E4, E5, E6\right]$ on side $\left[E9, E10, E11, E12, E13, E14\right]$ on the other and $\left[E7, E8\right]$ in an ambiguous middle position. Note that this is roughly the same partitioning obtained by \citep{doreian2004generalized} and reproduced by \citet{kovacs2010generalized} and \citet{lizardo2024two} using a generalized relational similarity approach. This indicates that persons and groups receive similar scores in the first dimension of CA only when they have similar connectivity similar to \textit{similar} groups, where group similarity is defined dually as having members in common who are themselves similar.

@@

\subsection{Normalized Similarity Plot}
To further appreciate the links between CA, reflective centralities, and community partitioning in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be thought of as giving the similarity between pairs of people weighted by the size of the groups they belong to (so that people are more similar when they share smaller memberships).\footnote{Note that \citet[eq.2]{newman2001scientific} proposes a slightly modified version of this two-mode similarity score for people---in Newman's case, scientists---except that the one mode projection is normalized as $A(Dg - I)^{-1}A^T$; namely, the size of the ``group''---the number of co-authors on a scientific paper---minus one. Substantively, this is unlikely to make much difference as the rank order of dyads by similarity between the two scores will be identical.} The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}, which is the similarity of groups weighted by the activity of the people in them (so that groups are more similar when they share members who do not belong to many other groups). As \citet{van2021correspondence} show, the eigenvectors corresponding to (the Laplacian\footnote{For the similarity matrix $S$, the Laplacian is defined as $D-S$ where $D$ is the matrix containing the degrees of either people or groups in the diagonal and zeroes in every other cell.} of) these similarity matrices are equivalent to $C^R$. This means that the relative spread of the eigenvalues of the degree-weighted similarity matrix provides information concerning the quality of the resulting community partition. These are shown as inset plots in Figure~\ref{fig:eigvec} for both people and groups. Note that in both cases, the first two eigenvalues separate from the rest, strongly indicating a dual-community structure in the Southern Women two-mode network.

@@

To fix the intuition, the top-left panel (a) of Figure~\ref{fig:ca-sim} shows a heatmap representation of the group-size weighted similarity matrix $S_p$ (for people); the top-right panel (b) shows the same plot for the activity-weighted similarity matrix $S_g$ (for groups). In both figures, darker squares indicate more similar node pairs, and closer to white indicate less similarity. Moreover, the rows of each matrix are sorted according to $C^R$ for both people and groups. As we can see, for both people and groups, the re-ordered matrix recovers maximally similar communities of nodes, where the similarity is based on their shared (degree-weighted) connections to entities in the other mode.

\subsection{CA Correspondence Plot}
The bottom panel of Figure~\ref{fig:ca-sim} shows the usual correspondence plot of the first two CA dimensions. People and groups are colored according to a six-cluster k-means solution based on the first nine eigenvectors of their respective similarity matrices. The plot reveals that the distances between nodes in the standard correspondence plot---usually taken to be a low-dimensional representation of the \textit{original} affiliation matrix \citep{borgatti1997network}---are best thought of as low-dimensional representations of the (other mode's) degree-normalized similarity network across people and groups. People with similar values in $S_p$---such as $\left[Olivia, Flora\right]$,  $\left[Katherine, Sylvia, Nora\right]$, $\left[Dorothy, Verne\right]$, and $\left[Brenda, Laura, Evelyn\right]$--- appear closer in the correspondence plot (and are assigned to the same cluster by the k-means algorithm). The same goes for events, those with similar values in $S_g$---such as $\left[E13, E14\right]$ and $\left[E1, E2, E3\right]$---appear closer in the correspondence plot, while those with dissimilar values are placed far apart. The distances between same-mode entities in the correspondence plot will thus be a function of their (inverse) degree-normalized similarity.

Note that this differs from the usual interpretation of the CA correspondence plot, which is usually taken to bring together nodes with ``similar'' connectivity patterns, where similarity is presumed to be a function of their raw row profiles (for people) or column profiles (for groups). This interpretation implies (for instance) that two people who attend many of the same events or two groups with many members will appear close in the plot. But this (still common) interpretation is off the mark. What the CA correspondence plot distance captures is, instead, people and groups that are \textit{surprisingly} similar (e.g., from the point of view of a suitable null model, like independence) after taking people's activity levels and the sizes of the groups they belong to into account.\footnote{Note that this, finding ``surprising'' similarities in terms of the indirect paths linking nodes in a network after the main effects of node-connectivity are considered, is the same rationale given by \cite{leicht2006vertex} for weighting the \citet{katz1953new} similarity matrix by the degrees of the corresponding nodes \citep[see][68, eqs. 2.13 and 2.14]{fouss2016algorithms}. As we have seen, this is precisely the key contrast between the Bonacich and the CA dual centrality measures for two-mode networks.} Thus, people who share memberships in small groups will be closer in the diagram than people who share memberships in big groups. In the same way, groups that share people with few memberships will be closer in the diagram than those sharing people with many other memberships.

@@

\subsection{One-Mode Projection Matrix and Eigevenctor Plot}
This ambiguity in interpretation may stem from the fact that CA is usually seen as a technique to generate a plot that provides a ``low-dimensional approximation to the input data'' \citep[125]{faust2005using}, where the ``input data'' is presumed to be the original affiliation matrix $\mathbf{A}$. But as we have seen, this is \textit{not} what CA is designed to do. Instead, CA is meant to provide a low-dimensional approximation of a \textit{transformed} version of the input data, where the transformation is meant to adjust for people's activity levels and group sizes. Notably, if a low-dimensional representation of the original ``input data'' ($\mathbf{A}$) is what we are after, this may be more closely approximated by the first two eigenvectors of the usual one-mode projections of the matrix (where the first is, of course, the standard Bonacich dual centrality score).
